---
title: "Airline data in the United States from 2017 to 2023"
date: 2023-09-15
output: html_notebook
---
Today, we make a representation from a specific airline/location about its passengers and the past 6 years.

```{r}
library(here)
library(nycflights13)
library(janitor)
library(tidyverse)
```

Now we load our files from the zip. This is from 2017 to May 2023. (https://www.transtats.bts.gov/DL_SelectFields.aspx?gnoyr_VQ=FIM&QO_fu146_anzr=Nv4+Pn44vr45).

```{r}
thisfile = here("data_raw", "DL_SelectFields_2017.zip")
df2017 = read_csv(thisfile) %>% clean_names()

thisfile = here("data_raw", "DL_SelectFields_2018.zip")
df2018 = read_csv(thisfile) %>% clean_names()

thisfile = here("data_raw", "DL_SelectFields_2019.zip")
df2019 = read_csv(thisfile) %>% clean_names()

thisfile = here("data_raw", "DL_SelectFields_2020.zip")
df2020 = read_csv(thisfile) %>% clean_names()

thisfile = here("data_raw", "DL_SelectFields_2021.zip")
df2021 = read_csv(thisfile) %>% clean_names()

thisfile = here("data_raw", "DL_SelectFields_2022.zip")
df2022 = read_csv(thisfile) %>% clean_names()

thisfile = here("data_raw", "DL_SelectFields_2023.zip")
df2023 = read_csv(thisfile) %>% clean_names()
```

Finally, we combine all the tables together and assign it to the variable df. Since we only care about the flights with passengers, we are going to filter that out.

```{r}
df = bind_rows(df2017, df2018, df2019, df2020, df2021, df2022, df2023)
df = df |> filter(passengers>0)
```

We want to graph the passengers and months to see which months of which year have what type of pattern.

```{r}
df2 = df |> filter(origin == "MIA" | dest == "MIA")
df2= df2 |> mutate(month_year = year+month/12)
ggplot(data = df2, mapping = aes(month_year)) + geom_histogram(aes(weight=passengers), color="black", fill="lightblue")
```

Question - do more folks travel "LGA to MIA" or "MIA to LGA"?

We created 2 data frames, one for LGA to MIA and another for MIA to LGA, mutated the data to be monthly, and then combined the data frames into one.

Then, we plotted the data and filtered by origin, so you could see both bar charts together on one graph.
```{r}
df3 = df |> filter(origin == "LGA" & dest == "MIA")
df4 = df |> filter(origin == "MIA" & dest == "LGA")

df3 = df3 |> mutate(month_year = year+month/12)
df4 = df4 |> mutate(month_year = year+month/12)

df3_4 = bind_rows(df3, df4)

ggplot(data = df3_4, mapping = aes(month_year)) + geom_bar(aes(weight=passengers, color=origin, fill=origin), position = 'dodge')

ggplot(data = df3_4, mapping = aes(month_year)) + geom_histogram(aes(weight=passengers), color="black", fill="lightblue")
```


