---
title: 'Experiment 1: Bike data'
author: "Naida Torres"
date: "2023-10-18"
output: html_notebook
---

##### Citation

We are continuing from Homework 1;

This experiment is largely borrowed from Coursework created Prof. Alexandra Chouldechova at Carnegie Mellon University.  I have modified slightly and am using in accordance with the Creative Commons License:
 
Creative Commons Attribution-NonCommercial-ShareAlike 4.0 International License(http://creativecommons.org/licenses/by-nc-sa/4.0/)

### Packages

```{r}
library(knitr)
library(tidyverse)
library(here)
library(broom)
library(janitor)
```

### Introduction: Bikeshare data

For this problem we'll be working with the data from a DC bike sharing app from September 2023 specifically.

### Data pre-processing 

Let's start by loading the data.

```{r}
thisfile = here("data_raw", "202309-capitalbikeshare-tripdata.zip")
df1 = read_csv(thisfile) %>% clean_names()
```

```{r}
df1 %>%
  slice_head(n = 100000) %>%
  ggplot(aes(start_lng, start_lat)) +
  geom_point()
```


